---
title: "R Notebook"
output: html_notebook
---
```{r}
library(readr)
library(readxl)
library(tidyr)
library(dplyr)
library(immunarch)
library(data.table)
```

Load subsampled data
```{r}
# this object was created in the immunarch_analysis script
sub <- readRDS(file="sub_BC.Rdata")
```

# Create dataframe and perform feature filtering

generate dataframe: features x observations. 
proportions of public clonotype sharing
```{r}
public_prop <- pubRep(sub, .quant = "prop", .verbose = F, .col = "aa")
```

```{r}
pub_prop_df = as.data.frame(public_prop)
pub_prop_df[is.na(pub_prop_df)] = 0
pub_prop_df_wos <- subset(pub_prop_df, select = -c(Samples) )
```

transpose dataframe: observations x features
```{r}
merged <- pub_prop_df_wos
merged_t <- setNames(data.frame(t(merged[,-1])), merged[,1])
```

turn data into matrix
```{r}
data_m <- data.matrix(merged_t)
```

select top 600 features
1. Count how many non zero cells in each column. (How many times does each feature appear?)
```{r}
samples_count <- colSums(data_m != 0)
```

2. indices of the sorted vector from high to low and keep top 600.
```{r}
feature_selection <- order(samples_count, decreasing = T)[1:600]
```

3. select those indices from data table.
```{r}
data_fs <- merged_t[,feature_selection]
```

export to csv
```{r}
write.csv(data_fs,"./data_600_ab.csv", row.names = T)
```

