---
title: "R Notebook"
output: html_notebook
---
```{r warning = FALSE}
library("immunarch")
library("purrr")
library("stringr")
library(forcats)
library(data.table)
library(ggpubr)
library(scales)
library(plotly)
library(dplyr)
```

All Samples (H+BC) TRA and TRB 
```{r}
#### BLOOD
immdata<-repLoad("/path/to/processed/data/")
saveRDS(immdata, "immdata_BC.Rdata")
```

Low clones
```{r}
clones<-repExplore(immdata$data, .method = "clones")
blood_order = arrange(clones, Clones)
rownames(blood_order) = c()
head(blood_order, 8)
```

# Subsample
1. Omit 21_H, 49_H, 15_BC, 11_BC
```{r}
immdata$data[which(names(immdata$data) %in% c("21_A_B_H", "49_A_B_H", "15_A_B_BC", "11_A_B_BC"))] <- NULL

```

2. subsample
```{r}
sub = repSample(immdata$data, .method = "downsample", .n = 24557)
```

```{r}
# this object will be used in the create_df script as well.
saveRDS(sub, "sub_BC.Rdata")
```

# Initial data analysis
```{r}
sub_h <- sub[grepl("H",names(sub))]
sub_bc <- sub[grepl("BC",names(sub))]
```

analysis after subsampling
sub_h and sub_bc contain both TRA and TRB

0. unique
1. top
2. rare
3. diversity
    a. gini
    b. True diversity "div"
    c. Inverse Simpson index "inv.simp"
    d. Gini-Simpson index "gini.simp"


0. unique
```{r}
uniq_h<-repExplore(sub_h, .method = "volume")

uniq_df_h <- data.frame(uniq_h)

plt_uniq_h <- ggplot(uniq_df_h,aes(x=Sample,y=Volume))+
  geom_bar(stat='identity',position='stack')+
  labs(title = "Unique clonotypes in Healthy Donors peripheral blood", x = "Sample", y = "Count") + 
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90))+
  theme(axis.text.x = element_blank(), panel.grid = element_blank())+
  scale_y_continuous(labels = comma)

plt_uniq_h
```

```{r}
uniq_bc<-repExplore(sub_bc, .method = "volume")

uniq_df_bc <- data.frame(uniq_bc)

plt_uniq_bc <- ggplot(uniq_df_bc,aes(x=Sample,y=Volume))+
  geom_bar(stat='identity')+
  labs(title = "Unique clonotypes in BC peripheral blood", x = "Sample", y = "Count") +
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90))+
  theme(axis.text.x = element_blank(),  panel.grid = element_blank())+
  scale_y_continuous(labels = comma)

plt_uniq_bc
```

t test
```{r}
uniq_y_h <- uniq_h$Volume
uniq_y_bc <- uniq_bc$Volume

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(uniq_y_h, uniq_y_bc)$p.value # 0.006
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(uniq_y_h, uniq_y_bc, var.equal = var.equal)

# p value = 0.003. significant difference!
```


1. top
```{r}
top_h <- repClonality(sub_h, .method = "top", .head= c(10,100,1000,3000,10000,30000)) 

top_df <- data.frame(top_h)
top_n <- mutate(top_df, "[11:100)" = X100 - X10, 
                  "[101:1000)" = X1000 - X100, 
                  "[1001:3000)" = X3000 - X1000, 
                  "[3001:10000)" = X10000 - X3000,
                  "[10001:30000)" = X30000 - X10000,
                  #"[30001:1e+05)" = X1e.05 - X30000,
                  #"[100001:3e+05)" = X3e.05 - X1e.05,
                  "Patient_ID" = rownames(top_df),
                  Num = as.numeric(map(str_split(Patient_ID, regex("_"),2),1)),
                  Patient_ID = fct_reorder(Patient_ID, Num))

top_n1 <- top_n[,c("Patient_ID","X10","[11:100)", "[101:1000)","[1001:3000)","[3001:10000)","[10001:30000)")] 
top_n1 = top_n1 %>% rename( "[1:10)" = "X10")

test <- reshape2::melt(top_n1)
plt <- ggplot(test,aes(x=Patient_ID,y=value,fill=variable))+
  geom_bar(stat='identity',position='stack')+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Top clonal proportion HD", x = "Sample", y = "Occupied repertoire space") + 
  guides(fill=guide_legend(title="Clonotype indices"))+
  scale_fill_brewer(palette = "RdYlBu")
plt
```

```{r}
top_bc <- repClonality(sub_bc, .method = "top", .head= c(10,100,1000,3000,10000,30000)) 

top_df_bc <- data.frame(top_bc)
top_n_bc <- mutate(top_df_bc, "[11:100)" = X100 - X10, 
                  "[101:1000)" = X1000 - X100, 
                  "[1001:3000)" = X3000 - X1000, 
                  "[3001:10000)" = X10000 - X3000,
                  "[10001:30000)" = X30000 - X10000,
                  #"[30001:1e+05)" = X1e.05 - X30000,
                  #"[100001:3e+05)" = X3e.05 - X1e.05,
                  "Patient_ID" = rownames(top_df_bc),
                  Num = as.numeric(ifelse(str_detect(Patient_ID, "^\\d_A_B_BC"), #
                                          map(str_split(Patient_ID, regex("_"),2),1), 
                                          substring(map(str_split(Patient_ID, regex("_"),2),1), 2)
                                          )
                                   ),
                  Patient_ID = fct_reorder(Patient_ID, Num)
                  )

top_n1_bc <- top_n_bc[,c("Patient_ID","X10","[11:100)", "[101:1000)","[1001:3000)","[3001:10000)","[10001:30000)")] # ,"[30001:1e+05)", "[100001:3e+05)"
top_n1_bc = top_n1_bc %>% rename( "[1:10)" = "X10")

test_bc <- reshape2::melt(top_n1_bc)
plt_bc <- ggplot(test_bc,aes(x=Patient_ID,y=value,fill=variable))+
  geom_bar(stat='identity',position='stack')+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Top clonal proportion BC", x = "Sample", y = "Occupied repertoire space") + 
  guides(fill=guide_legend(title="Clonotype indices"))+
  scale_fill_brewer(palette = "RdYlBu")
plt_bc
```

t test
```{r}
top_y_h <- top_n1$`[1:10)`
top_y_bc <- top_n1_bc$`[1:10)`

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(top_y_h, top_y_bc)$p.value # 0.027
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(top_y_h, top_y_bc, var.equal = var.equal)

# p-value = 0.24. no significant difference
```


2. rare
```{r}
rare_h <- repClonality(sub_h, .method = "rare") #%>% vis()

rare_df_h <- data.frame(rare_h)
rare_h_n <- mutate(rare_df_h, "2 - 3" = X3 - X1, 
                  "4 - 10" = X10 - X3, 
                  "11 - 30" = X30 - X10, 
                  "31 - 100" = X100 - X30,
                  "101 - MAX" = MAX - X100,
                  "Patient_ID" = rownames(rare_df_h),
                  Num = as.numeric(map(str_split(Patient_ID, regex("_"),2),1)),
                  Patient_ID = fct_reorder(Patient_ID, Num))
rare_h_n1 <- rare_h_n[,c("Patient_ID","X1","2 - 3", "4 - 10", "11 - 30","31 - 100","101 - MAX")]
rare_h_n1 = rare_h_n1 %>% rename( "1" = "X1")
test_rare_h <- reshape2::melt(rare_h_n1)

plt_rare_h <- ggplot(test_rare_h,aes(x=Patient_ID,y=value,fill=variable))+
  geom_bar(stat='identity',position='stack')+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Rare clonal proportion HD", x = "Sample", y = "Occupied repertoire space") + 
  guides(fill=guide_legend(title="Clonotype counts"))+
  scale_fill_brewer(palette = "RdYlBu")

plt_rare_h
```

```{r}
rare_bc <- repClonality(sub_bc, .method = "rare") #%>% vis()

rare_df_bc <- data.frame(rare_bc)
rare_n_bc <- mutate(rare_df_bc, "2 - 3" = X3 - X1, 
                  "4 - 10" = X10 - X3, 
                  "11 - 30" = X30 - X10, 
                  "31 - 100" = X100 - X30,
                  "101 - MAX" = MAX - X100,
                  "Patient_ID" = rownames(rare_df_bc),
                  Num = as.numeric(ifelse(str_detect(Patient_ID, "^\\d_A_B_BC"), #
                                          map(str_split(Patient_ID, regex("_"),2),1), 
                                          substring(map(str_split(Patient_ID, regex("_"),2),1), 2)
                                          )
                                   ),
                  Patient_ID = fct_reorder(Patient_ID, Num))
rare_n1_bc <- rare_n_bc[,c("Patient_ID","X1","2 - 3", "4 - 10", "11 - 30","31 - 100","101 - MAX")]
rare_n1_bc = rare_n1_bc %>% rename( "1" = "X1")
test_rare_bc <- reshape2::melt(rare_n1_bc)

plt_rare_bc <- ggplot(test_rare_bc,aes(x=Patient_ID,y=value,fill=variable))+
  geom_bar(stat='identity',position='stack')+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), legend.position = "none", complete = F, panel.grid = element_blank())+
  labs(title = "Rare clonal proportion BC", x = "Sample", y = "Occupied repertoire space") + 
  guides(fill=guide_legend(title="Clonotype counts"))+
  scale_fill_brewer(palette = "RdYlBu")
plt_rare_bc
```

t test
```{r}
rare_y_h = rare_h_n1[,2]

rare_y_bc = rare_n1_bc[,2]

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(rare_y_h, rare_y_bc)$p.value # 0.012 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(rare_y_h, rare_y_bc, var.equal = var.equal)

# p-value = 0.001, significant difference!
```


3. diversity- gini
```{r}
gini_h <- repDiversity(sub_h, .method = "gini", .verbose = F)

gini_df_h = data.frame(gini_h)
gini_df_h$sample <- rownames(gini_df_h)
gini_df_h$Num = as.numeric(map(str_split(gini_df_h$sample, regex("_"),2),1))
gini_df_h$Patient_ID = fct_reorder(gini_df_h$sample, gini_df_h$Num)

plt_gini_h<- ggplot(gini_df_h, aes(Patient_ID, gini_h)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Gini Diversity HD", x = "Sample", y = "gini index")+
  theme(legend.position = "none")

plt_gini_h
```

```{r}
gini_bc <- repDiversity(sub_bc, .method = "gini", .verbose = F)

gini_df_bc = data.frame(gini_bc)
gini_df_bc$sample <- rownames(gini_df_bc)
gini_df_bc$Num = as.numeric(ifelse(str_detect(gini_df_bc$sample, "^\\d_A_B_BC"), #
                                          map(str_split(gini_df_bc$sample, regex("_"),2),1), 
                                          substring(map(str_split(gini_df_bc$sample, regex("_"),2),1), 2)
                                          )
                                   )
gini_df_bc$Patient_ID = fct_reorder(gini_df_bc$sample, gini_df_bc$Num)

plt_gini_bc<- ggplot(gini_df_bc, aes(Patient_ID, gini_bc)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Gini Diversity BC", x = "Sample", y = "gini index")+
  theme(legend.position = "none")

plt_gini_bc
```

t test
```{r}
gini_y_h = gini_df_h$gini_h

gini_y_bc = gini_df_bc$gini_bc

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(gini_y_h, gini_y_bc)$p.value # 0.005 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(gini_y_h, gini_y_bc, var.equal = var.equal)

# p-value = 0.005,  significant difference!
```


3b. diversity - True diversity "div"
```{r}
true_d_h <- repDiversity(sub_h, .method = "div", .verbose = F)

true_d_h$Num = as.numeric(map(str_split(true_d_h$Sample, regex("_"),2),1))
true_d_h$Patient_ID = fct_reorder(true_d_h$Sample, true_d_h$Num)

plt_trued_h<- ggplot(true_d_h, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "True Diversity HD", x = "Sample", y = "True Diversity")+
  theme(legend.position = "none")

plt_trued_h
```

```{r}
true_d_bc <- repDiversity(sub_bc, .method = "div", .verbose = F)

true_d_bc$Num = as.numeric(ifelse(str_detect(true_d_bc$Sample, "^\\d_A_B_BC"), #
                                          map(str_split(true_d_bc$Sample, regex("_"),2),1), 
                                          substring(map(str_split(true_d_bc$Sample, regex("_"),2),1), 2)
                                          )
                                   )
true_d_bc$Patient_ID = fct_reorder(true_d_bc$Sample, true_d_bc$Num)

plt_trued_bc<- ggplot(true_d_bc, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "True Diversity BC", x = "Sample", y = "True Diversity")+
  theme(legend.position = "none")

plt_trued_bc
```

t test
```{r}
trued_y_h = true_d_h$Value

trued_y_bc = true_d_bc$Value

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(trued_y_h, trued_y_bc)$p.value 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(gini_y_h, gini_y_bc, var.equal = var.equal)

# p-value = 0.005, significant difference
```

3c. diversity - Inverse Simpson index "inv.simp"
```{r}
inv_h <- repDiversity(sub_h, .method = "inv.simp", .verbose = F)

inv_h$Num = as.numeric(map(str_split(inv_h$Sample, regex("_"),2),1))
inv_h$Patient_ID = fct_reorder(inv_h$Sample, inv_h$Num)

plt_inv_h<- ggplot(inv_h, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Inverse Simpson Index Diversity HD", x = "Sample", y = "Inverse Simpson index")+
  theme(legend.position = "none")

plt_inv_h
```

```{r}
inv_bc <- repDiversity(sub_bc, .method = "inv.simp", .verbose = F)

inv_bc$Num = as.numeric(ifelse(str_detect(inv_bc$Sample, "^\\d_A_B_BC"), #
                                          map(str_split(inv_bc$Sample, regex("_"),2),1), 
                                          substring(map(str_split(inv_bc$Sample, regex("_"),2),1), 2)
                                          )
                                   )
inv_bc$Patient_ID = fct_reorder(inv_bc$Sample, inv_bc$Num)

plt_inv_bc<- ggplot(inv_bc, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Inverse Simpson Index Diversity BC", x = "Sample", y = "Inverse Simpson index")+
  theme(legend.position = "none")

plt_inv_bc
```

t test
```{r}
inv_y_h = inv_h$Value

inv_y_bc = inv_bc$Value

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(inv_y_h, inv_y_bc)$p.value 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(inv_y_h, inv_y_bc, var.equal = var.equal)

# p-value = 0.0007, significant difference!!
```


3d. diversity - Gini-Simpson index "gini.simp"
```{r}
simp_h <- repDiversity(sub_h, .method = "gini.simp", .verbose = F)

simp_h$Num = as.numeric(map(str_split(simp_h$Sample, regex("_"),2),1))
simp_h$Patient_ID = fct_reorder(simp_h$Sample, simp_h$Num)

plt_simp_h<- ggplot(simp_h, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Gini Simpson index Diversity HD", x = "Sample", y = "Gini-Simpson index")+
  theme(legend.position = "none")

plt_simp_h
```

```{r}
simp_bc <- repDiversity(sub_bc, .method = "gini.simp", .verbose = F)

simp_bc$Num = as.numeric(ifelse(str_detect(simp_bc$Sample, "^\\d_A_B_BC"), #
                                          map(str_split(simp_bc$Sample, regex("_"),2),1), 
                                          substring(map(str_split(simp_bc$Sample, regex("_"),2),1), 2)
                                          )
                                   )
simp_bc$Patient_ID = fct_reorder(simp_bc$Sample, simp_bc$Num)

plt_simp_bc<- ggplot(simp_bc, aes(Patient_ID, Value)) + #, fill = sample
  geom_col()+
  theme_bw()+
  #theme(axis.text.x = element_text(angle = 90), complete = F, panel.grid = element_blank())+
  theme(axis.text.x = element_blank(), complete = F, panel.grid = element_blank())+
  labs(title = "Gini Simpson index Diversity BC", x = "Sample", y = "Gini-Simpson index")+
  theme(legend.position = "none")

plt_simp_bc
```

t test
```{r}
simp_y_h = simp_h$Value

simp_y_bc = simp_bc$Value

# p-val < 0.05 --> var.equal = F 
# else var.equal = T
p.var <- var.test(simp_y_h, simp_y_bc)$p.value # 0.000001 
var.equal <- p.var >= 0.05

# p-val < 0.05 --> significant difference
# else no sig dif
t.test(simp_y_h, simp_y_bc, var.equal = var.equal)

# p-value = 0.65, no significant difference
```
